BERT Model Compression and Weight Sharing

### 2019-05-13 paper/paper.tex
Compression ratio discussion in results

Deep Neural Networks (DDNs) have become the state-of-art for many tasks in
language modeling. At the same time, they have become progressively larger in
size, due to both larger out-of-box pre-trained embeddings and deeper
networks. However, in order to deploy the models on more resource-constrained
devices, such as mobile devices or even older personal laptops, as opposed to
on cloud virtual machines, it is crucial to reduce the size of the models.

@@

The most important contribution of this paper is that we proposed a new
weight-sharing method called ``Difference Pruning'' which imposes constraint on
the sparcity of weight difference with the pre-trained BERT model. While
uncompressed fine-tuned BERT model achieves 92.9\% validation accuracy on
SST-2, the Difference Pruning model that keeps 99\% of the pre-trained BERT
weights achieves 91.3\% on the validation set. This enables multiple language
modeling tasks to share the weights of pre-trained BERT and each incurs minimal
extra storage space requirements.

@@

An additional benefit from knowledge distillation arises from BERT's
pre-training. Because BERT is pre-trained on large amounts of unstructured
data, it contains a powerful language model which likely would be impossible
to learn from the specific task's data-set. To an extent, relevant features
from this language model are what make BERT so powerful once it's fine-tuned.
Training the student network on BERT's output thus allows some of this
``outside knowledge'' to percolate toward the student network.

@@

  Pruning is an out-of-the-box procedure that works with any models. However,
  as a price of the generality, pruning fails to be tailored to the nature of
  BERT being a pre-trained model. Suppose the pretrained weights $\bphi_0 $
  capture certain structure of language, then dramatic distortions to
  $\bphi_0$ essentially discards the information, especially early in the
  training loop. We may partially remedy this issue by either using a dynamic
  pruning rate, where pruning rate is small early in the training loop or by
  using layer-specific pruning rates, with layers in $\bert$ having a lower
  pruning rate than layers in $G$.

@@

We distort via \[
\rankreduce(\bvartheta; r) = \tilde U \tilde \Sigma \tilde V^\top,
\]
where $\tilde U, \tilde \Sigma, \tilde V$ are conforming matrices such that the
smallest values on the spectrum $\Sigma$ is zeroed out, as well as corresponding
values in $U,V$. If we keep only the $r$ largest spectral values, then we store
$\tilde U \tilde \Sigma$, an $n\times r$ matrix, and $\tilde V$, an $m\times r$
matrix. Thus, instead of storing $2nm$ floats, we may store $r(n+m)$, achieving
a compression rate of \[
\rho = \frac{r(n+m)}{2mn}.
\]
Note however, that storing matrices in this manner requires an additional
matrix multiplication in evaluation time, since we have to compute the product
$\tilde U \tilde \Sigma \tilde V$.

@@

@@

Our most striking result is the robust performance of difference pruning. Even
keeping upwards of 99\% of the weights unchanged gives us performance that is
very competitive with the unrestricted model. Most of our experiments are on
SST-2, but we see success with this method on QQP as well, suggesting that
small changes to pre-trained BERT are sufficient for performance on a specific
task.

>>>>

Deep Neural Networks (DNNs) have become the state-of-art for many tasks in
language modeling. At the same time, they have become progressively larger in
size, due to both larger out-of-box pretrained embeddings and deeper
networks. However, in order to deploy the models on more resource-constrained
devices, such as mobile devices or even older personal laptops, as opposed to
on cloud virtual machines, it is crucial to reduce the size of the models.

@@

The most important contribution of this paper is that we proposed a new
weight-sharing method called ``Difference Pruning'' which imposes constraint on
the sparcity of weight difference with the pretrained BERT model. While
uncompressed fine-tuned BERT model achieves 92.9\% validation accuracy on
SST-2, the Difference Pruning model that keeps 99\% of the pretrained BERT
weights achieves 91.3\% on the validation set. This enables multiple language
modeling tasks to share the weights of pretrained BERT and each incurs minimal
extra storage space requirements.

@@

An additional benefit from knowledge distillation arises from BERT's
pre-training. Because BERT is pretrained on large amounts of unstructured
data, it contains a powerful language model which likely would be impossible
to learn from the specific task's data-set. To an extent, relevant features
from this language model are what make BERT so powerful once it's fine-tuned.
Training the student network on BERT's output thus allows some of this
``outside knowledge'' to percolate toward the student network.

@@

  Pruning is an out-of-the-box procedure that works with any models. However,
  as a price of the generality, pruning fails to be tailored to the nature of
  BERT being a pretrained model. Suppose the pretrained weights $\bphi_0 $
  capture certain structure of language, then dramatic distortions to
  $\bphi_0$ essentially discards the information, especially early in the
  training loop. We may partially remedy this issue by either using a dynamic
  pruning rate, where pruning rate is small early in the training loop or by
  using layer-specific pruning rates, with layers in $\bert$ having a lower
  pruning rate than layers in $G$.

@@

We distort via \[
\rankreduce(\bvartheta; r) = \tilde U \tilde \Sigma \tilde V^\top,
\]
where $\tilde U, \tilde \Sigma, \tilde V$ are conforming matrices such that the
smallest values on the spectrum $\Sigma$ is zeroed out, as well as corresponding
values in $U,V$. If we keep only the $r$ largest spectral values, then we store
$\tilde U \tilde \Sigma$, an $n\times r$ matrix, and $\tilde V$, an $m\times r$
matrix. Thus, instead of storing $2nm$ floats, we may store $r(n+m)$, achieving
a compression rate of \[
\rho = 1-\frac{r(n+m)}{2mn}.
\]
Note however, that storing matrices in this manner requires an additional
matrix multiplication in evaluation time, since we have to compute the product
$\tilde U \tilde \Sigma \tilde V$.

@@

Pruning admits a simple upper bound for compression, which is simply the
pruning ratio.\footnote{The exact compression rate is complicated, due to
the sparse data structure implemented for the compressed model} Compression
rate for difference pruning depends on the number of tasks the user
desires to perform, since difference pruning is understood to be a form of
weight sharing. In that context, the (upper bound) compression rate of
difference
pruning
is also self-evident. The compression rate achieved by rank reduction is \[
\rho = 1 - \frac{r(n+m)}{2mn}.
\]
Since most matrices in \texttt{BERT-base-uncased} is $768\times 768,$ the
compression rate is approximately \[
\rho \approx 1 - \frac{r}{768} = (74.0\%,
80.5\%,
87.0\%,
93.5\%)
\]
for $r = (200,150,100,50)$, respectively.

@@

Our most striking result is the robust performance of difference pruning. Even
keeping upwards of 99\% of the weights unchanged gives us performance that is
very competitive with the unrestricted model. Most of our experiments are on
SST-2, but we see success with this method on QQP as well, suggesting that
small changes to pretrained BERT are sufficient for performance on a specific
task.
